---
title: "finalAnalyses"
author: "Or Duek"
date: "6/7/2019"
output:
  html_document: default
  pdf_document: default
---
This is the final analysis for the VA paper. 
All changes will be done on this one

```{r loading_data,cache=FALSE, echo=FALSE}
# load data
library(tidyverse)
library(jtools)
library(survival)
library(survminer)
library(sjPlot)
vaDat <- read_csv('/home/or/Dropbox/postdoc/Yale/va_data/MED_Drop.csv')

vaDatclean <- vaDat
# remove more than 365 psychotherapies
vaDatclean$MHTOTALLc <- ifelse((vaDatclean$MHTOTALL > 365), (vaDatclean$MHTOTALLc = 365), (vaDatclean$MHTOTALLc = vaDatclean$MHTOTALL))

# set index date
a = '01/01/1960'
vaDatclean$PTSDINDr = as.Date(as.character(a), format = '%d/%m/%Y') + vaDatclean$PTSDIND
# Now take only patients who recieved medications. 
onlyMed_dat <- subset(vaDatclean, vaDatclean$PMEDRX==1)
# set number of days from index day to drop out
vaDatclean$totalDays <- as.Date(as.character(vaDatclean$MHDROPD),format = '%m/%d/%Y')  - as.Date(vaDatclean$PTSDINDr,format = '%Y/%m/%d')
head(vaDatclean)
```


```{r creating_vars, echo=FALSE}
# create a few variables before stratification
# Race
vaDatclean$race <- dplyr::case_when(vaDatclean$BLACK==1 ~ "Black",vaDatclean$WHITE==1 ~ "White", vaDatclean$HISPANIC==1 ~ "Hispanic", vaDatclean$ASIAN ==1 ~ "Asian", TRUE ~ "Other")
vaDatclean$race <- factor(vaDatclean$race)
# change levels white will be first
vaDatclean$race <- relevel(vaDatclean$race, "White")

# turn age to categorical
vaDatclean$ageCat <- cut(vaDatclean$AGE_OCT01, breaks = c(15, 25, 35, 45, 60, 90), labels = c("18-25", "26-35", "36-45", "45-60", "60+"))
plot(vaDatclean$ageCat)
# check comorbidity
vaDatclean$sumComorbid <- vaDatclean$DEMENT_ALZ + vaDatclean$ALC + vaDatclean$DRUG + vaDatclean$SCHIZOPHRENIA + vaDatclean$OTHERPSYCHOSIS + vaDatclean$BIPOLAR + vaDatclean$MAJORAFFECT + vaDatclean$OTHERDEP + vaDatclean$ANXIETY + vaDatclean$ADJUSTMENT + vaDatclean$PERSONALITY + vaDatclean$BORDERLINE

vaDatclean$MARRIED <- factor(vaDatclean$MARRIED)
vaDatclean$MARRIED <- relevel(vaDatclean$MARRIED, "1") # change reference level to married

vaDatclean$FEMALE <- factor(vaDatclean$FEMALE)
vaDatclean$VATXM <- factor(vaDatclean$VATXM) #, labels = c("Yes", "No"))
# Creating Psychotherapy 0 - none or 1 = exist
vaDatclean$PSYCTR <- dplyr::case_when(vaDatclean$INDPSY >1 | vaDatclean$GRPPSY >1 | vaDatclean$FAMPSY >1 ~ "yes", TRUE ~ "no")
vaDatclean$PSYCTR <- factor(vaDatclean$PSYCTR)
vaDatclean$PSYCTR <- relevel(vaDatclean$PSYCTR, "yes")

vaDatclean$sumComorbidfactor <- cut(vaDatclean$sumComorbid, breaks = c(-1,0,1,3,15), labels = c("0","1", "2-3","3+"))
plot(vaDatclean$sumComorbidfactor)

vaDatclean$ANXfactor <- factor(vaDatclean$ANXIETY)
vaDatclean$URBAN_RURAL <- factor(vaDatclean$URBAN_RURAL)
vaDatclean$MAJORAFFECT <- factor(vaDatclean$MAJORAFFECT)


vaDatclean$OEFOIF <- factor(vaDatclean$OEFOIF)
vaDatclean$OEFOIF <- relevel(vaDatclean$OEFOIF, "1")

vaDatclean$ADRXN_true <- case_when(vaDatclean$ADRXN > 0 ~ 1, TRUE ~0)
vaDatclean$APSRXN_true <- case_when(vaDatclean$APSRXN >0 ~ 1, TRUE ~0)
vaDatclean$AXSEDHPRXN_true <- case_when(vaDatclean$AXSEDHPRXN >0 ~ 1, TRUE ~0)
vaDatclean$STIMRXN_true <- case_when(vaDatclean$STIMRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$ACVMSRXN_true <- case_when(vaDatclean$ACVMSRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$LIRXN_true <- case_when(vaDatclean$LIRXN > 0 ~ 1, TRUE ~ 0)

vaDatclean$sumMed <- vaDatclean$LIRXN_true + vaDatclean$ACVMSRXN_true + vaDatclean$STIMRXN_true + vaDatclean$AXSEDHPRXN_true + vaDatclean$APSRXN_true + vaDatclean$ADRXN_true 
vaDatclean$sumMedCat <- factor(case_when(vaDatclean$sumMed==1 ~ '1', vaDatclean$sumMed == 2 ~ "2", vaDatclean$sumMed >= 3 ~ "3+"))

# create a censore variable who survived more than 30 days
vaDatclean$MONTH1 <- case_when(vaDatclean$totalDays <= 30 ~ 1, TRUE ~ 0)

# create comorbidity variable with MAJORAFFECT, OTHERDEP, ANXIETY, SCHIZOPHRENIA, OTHERPSYCHOSIS
vaDatclean$threeComb <- case_when(vaDatclean$MAJORAFFECT==1 ~ "Dep", vaDatclean$OTHERDEP ==1 ~ 'Dep', vaDatclean$ANXIETY==1 ~ 'Anx', vaDatclean$SCHIZOPHRENIA==1 ~ 'Psychosis', vaDatclean$OTHERPSYCHOSIS==1 ~ 'Psychosis', TRUE ~ 'noComb')
table(vaDatclean$threeComb)
vaDatclean$threeComb <- relevel(factor(vaDatclean$threeComb), 'noComb')



```



Lets take just medicated patients, before stratify to different groups of medication.

```{r filter_med, echo=FALSE}
# take just medicated patients. 
medicatedVa_dat <- filter(vaDatclean, PMEDRX==1)
head(medicatedVa_dat)
# take only patients with first MH visit after index date
medicatedVa_datFirstV <-  subset(medicatedVa_dat, !is.na(medicatedVa_dat$MHFUFRST))
medicatedVa_datFirstV$timeTofV <- factor(case_when(medicatedVa_datFirstV$MHFUFRST<=7 ~ "7", medicatedVa_datFirstV$MHFUFRST>7 & medicatedVa_datFirstV$MHFUFRST<=30 ~ "7-30", TRUE ~ "30+"))

medicatedVa_datFirstV$timeTofV <- relevel(medicatedVa_datFirstV$timeTofV, "30+" )
```
```{r time_censore_vars, echo=FALSE}
########## Create new censore and time variables (according to discussion with Ilan) #############################
# Build censor var for antidepressants
# will calculate ADDAY again - just to make sure we're ok
# as.Date(as.character(vaDatclean$MHDROPD),format = '%m/%d/%Y')  - as.Date(vaDatclean$PTSDINDr,format = '%Y/%m/%d')
# building a censoring variable by 180days (as we're not sure about the 365 one)
medicatedVa_datFirstV$antiDPCensore <- case_when(medicatedVa_datFirstV$ADDAY>=180 ~ 0, medicatedVa_datFirstV$ADDAY<180 ~ 1, is.na(medicatedVa_datFirstV$ADDAY) ~ NA_real_, TRUE ~ 3)
summary(medicatedVa_datFirstV$antiDPCensore)
table(medicatedVa_datFirstV$antiDPCensore)
table(medicatedVa_datFirstV$MHDROP)
# build cencor var for anxiolytics
medicatedVa_datFirstV$antiAXCensore <- case_when(medicatedVa_datFirstV$AXDAY>=180 ~ 0, medicatedVa_datFirstV$AXDAY<180 ~ 1, is.na(medicatedVa_datFirstV$AXDAY) ~ NA_real_, TRUE ~ 3)
table(medicatedVa_datFirstV$antiAXCensore)
# build cencor var for anti psychotic
medicatedVa_datFirstV$antiPCensore <- case_when(medicatedVa_datFirstV$APSDAY>=180 ~ 0, medicatedVa_datFirstV$APSDAY<180 ~ 1, is.na(medicatedVa_datFirstV$APSDAY) ~ NA_real_, TRUE ~ 3)
table(medicatedVa_datFirstV$antiPCensore)
table(medicatedVa_datFirstV$threeComb)
# build timevar for anti depressant (last -first rx)

# build timevar for anxiolytic (last-first rx)
# build timevar for anti psychotic (last - first rx)
# build total timevar (last rx - first rx between all those meds)
  # take date of first rx of first med and last rx of last med and calculate number of days
  # create var of first date (that compared antiDP, antiolytic and antipsychotic)
  # create a var of last date
  # new var of total days
# build total cencor var (three med)
 

##################################################################################################################
```

### When looking at total PTSD population, 224060 are without comorbidity and 234,673 are with depression
### When looking at medicated patients 

## Now we go medication by medication:
### First -- Anti depression 82,288 are without comorbidity and 194,145 with dep. 

```{r antiDP_analysis, echo=FALSE}
# now we clean data
# take only anti depressant without any medication before index day
antiDP_dat <- dplyr::filter(medicatedVa_datFirstV, !is.na(ADRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 
head(antiDP_dat)
summary(antiDP_dat$ADDAY)
sd(antiDP_dat$ADDAY)
# how many dropout within 180 days
table(antiDP_dat$antiDPCensore)
table(antiDP_dat$threeComb)
## survival model
surv_antiDP <- Surv(time = antiDP_dat$ADDAY, event = antiDP_dat$antiDPCensore)  
summary(surv_antiDP)
# now cox analysis
fit.coxpDP <- coxph(surv_antiDP ~ PSYCTR + MARRIED + OEFOIF + race + threeComb + FEMALE +sumComorbid + sumMed, 
                   data = antiDP_dat)
ggforest(fit.coxpDP, data = antiDP_dat)
summary(fit.coxpDP)
```
## Anxiolytics 
```{r anxiolytic_analysis, echo=FALSE}
# take only anxiolytics
antiANX_dat <- dplyr::filter(medicatedVa_datFirstV, !is.na(AXSEDHPRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 
head(antiANX_dat)
summary(antiDP_dat$AXDAY)
sd(antiDP_dat$AXDAY, na.rm = TRUE)
# how many dropout at 180 days
table(antiANX_dat$antiAXCensore)
## survival model
surv_antiANX <- Surv(time = antiANX_dat$AXDAY, event = antiANX_dat$antiAXCensore)  
summary(surv_antiANX)
# now cox analysis
fit.coxpANX <- coxph(surv_antiANX ~ PSYCTR + MARRIED + OEFOIF + race +threeComb + FEMALE+ sumComorbid + sumMed, 
                   data = antiANX_dat)
ggforest(fit.coxpANX, data = antiANX_dat)
summary(fit.coxpANX)
```
## Antipsychotics
```{r antiPsychotic_analysis, echo=FALSE}
antiAPX_dat <- dplyr::filter(medicatedVa_datFirstV, !is.na(APSRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 
# how many drop out at 180 days
table(antiAPX_dat$antiPCensore)
## Check mean and std
summary(antiDP_dat$APSDAY)
sd(antiDP_dat$APSDAY, na.rm = TRUE)
## survival model
surv_antiAP <- Surv(time = antiAPX_dat$APSDAY, event = antiAPX_dat$antiPCensore)  
summary(surv_antiAP)
# now cox analysis
fit.coxpAP <- coxph(surv_antiAP ~ PSYCTR + MARRIED + OEFOIF + race +threeComb + FEMALE+ sumComorbid + sumMed, 
                   data = antiAPX_dat)
ggforest(fit.coxpAP, data = antiAPX_dat)
summary(fit.coxpAP)
```

## All 3 types of meds toghether:
```{r allThree_analysis, echo=FALSE}
# Take anti depressants, anti psychotics and anxiolytics without any prior medications. 
threeMeds_dat <- filter(medicatedVa_datFirstV, is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX), !is.na(AXSEDHPRXN) | !is.na(ADRXN) | !is.na(APSRXN))
head(threeMeds_dat)
  #filter(medicatedVa_dat, !is.na(APSRXN) | !is.na(AXSEDHPRXN) | !is.na(ADRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX))
# build total timevar (last rx - first rx between all those meds)
  # take date of first rx of first med and last rx of last med and calculate number of days
  # create var of first date (that compared antiDP, antiolytic and antipsychotic)
threeMeds_dat$firstDateMed <- pmin(as.Date(threeMeds_dat$ADFSTRX,format = '%m/%d/%Y' ), as.Date(threeMeds_dat$APSFSTRX, format = '%m/%d/%Y'), as.Date(threeMeds_dat$AXFSTRX, format = '%m/%d/%Y'), na.rm = TRUE)
  # create a var of last date
threeMeds_dat$lastDateMed <- pmax(as.Date(threeMeds_dat$ADLSTRX,format = '%m/%d/%Y' ), as.Date(threeMeds_dat$APSLSTRX, format = '%m/%d/%Y'), as.Date(threeMeds_dat$AXLSTRX, format = '%m/%d/%Y'), na.rm = TRUE)
# new var of total days
threeMeds_dat$totalDaysThreeMeds <- as.Date(threeMeds_dat$lastDateMed, format = '%m/%d/%Y') - as.Date(threeMeds_dat$firstDateMed, format = '%m/%d/%Y')
hist(as.numeric(threeMeds_dat$totalDaysThreeMeds))
#ggplot(threeMeds_dat, aes(x=totalDaysThreeMeds)) + geom_histogram(fill = "Orange") + theme_minimal() + xlab("Number of days with medication") + ylab("Number of Patients") + geom_line(aes(x=totalDaysThreeMeds,y=cumsum(totalDaysThreeMeds)), col="red", lwd=1)
# build total cencor var (three med)
threeMeds_dat$threeMedsCensore <- case_when(threeMeds_dat$totalDaysThreeMeds>=180 ~ 0, threeMeds_dat$totalDaysThreeMeds<180 ~ 1, is.na(threeMeds_dat$totalDaysThreeMeds) ~ NA_real_, TRUE ~ 3)
table(threeMeds_dat$threeMedsCensore)

## now we do survival analysis
surv_ThreeMeds <- Surv(time = threeMeds_dat$totalDaysThreeMeds, event = threeMeds_dat$threeMedsCensore)  
summary(surv_ThreeMeds)
# now cox analysis
fit.coxpThreeMeds <- coxph(surv_ThreeMeds ~ PSYCTR + MARRIED + OEFOIF + race + threeComb + FEMALE + sumComorbid + sumMed, 
                   data = threeMeds_dat)
ggforest(fit.coxpThreeMeds, data = threeMeds_dat)
summary(fit.coxpThreeMeds)
```
## Results are the same in three medications toghether.
## Sample caracteristics
```{r sample_Car}
hist(threeMeds_dat$AGE_OCT01)
mean(threeMeds_dat$AGE_OCT01, na.rm = TRUE)
sd(threeMeds_dat$AGE_OCT01, na.rm = TRUE)
summary(threeMeds_dat$sumComorbidfactor)
mean(threeMeds_dat$totalDaysThreeMeds, na.rm = TRUE)
sd(threeMeds_dat$totalDaysThreeMeds)
```

# 30 days:
Doing a binomial regression model with censore =30 as dependent variable. AGE, number of psychotherapy sessions marital status as dependent, controlling for number of medications and comorbidity index. 
```{r 30Days_binomial}

threeMeds_dat$censore30 <- case_when(threeMeds_dat$totalDaysThreeMeds>=30 ~ 0, threeMeds_dat$totalDaysThreeMeds<30 ~ 1, is.na(threeMeds_dat$totalDaysThreeMeds) ~ NA_real_)

model1 <- glm(censore30 ~ AGE_OCT01 + PSYCTHR + MARRIED + race + sumComorbid+ sumMed + threeComb, data=threeMeds_dat, family = binomial)
summary(model1)
model1_summ <- as.data.frame(summ(model1, exp = TRUE, scale =TRUE, digits = 5)$coeftable, optional = TRUE)  #just taking coeficient table and creating data frame to plot 
```
```{r PlotBinom}
# plott the above
print(plot_summs(model1, scale = TRUE, exp = TRUE, colors = "Qual1", 
           coefs = c("Age" = "AGE_OCT01", "# Psych Sessions" = "PSYCTHR", "Asian" = "raceAsian", "Black" = "raceBlack", "Hispanic" = "raceHispanic", "Other" = "raceOther",
                     "Not Married" = "MARRIED", "Comorbidity Index" = "sumComorbid", "# of Medications" = "sumMed", "AnxMed" = "threeCombAnx", "DepMed" = "threeCombDep", "PsycMed" = "threeCombPsychosis")))
```


```{r}
threeMeds_30days <- filter(threeMeds_dat, threeMeds_dat$totalDaysThreeMeds<=30)
53475/154953
threeMeds_30days$censore30 <- case_when(threeMeds_30days$totalDaysThreeMeds>=30 ~ 0, threeMeds_30days$totalDaysThreeMeds<30 ~ 1, is.na(threeMeds_30days$totalDaysThreeMeds) ~ NA_real_)
table(threeMeds_30days$censore30)
## now we do survival analysis
surv_30days <- Surv(time = threeMeds_dat$totalDaysThreeMeds, event = threeMeds_dat$censore30)  
summary(surv_30days)
# now cox analysis
fit.coxp30days <- coxph(surv_30days ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbid + sumMed, 
                   data = threeMeds_dat)
ggforest(fit.coxp30days, data = threeMeds_30days)
summary(fit.coxp30days)
```


## Combine PCL scores with 30 days data
```{r pcl_charac}
# characteristics of pcl sample
pclNoNa <- filter(threeMeds_dat, !is.na(BPCLTOT))
# pcl database contains 
nrow(pclNoNa) 


### A total of 24,418 patients had PCL scores.

# create five factor pcl scores
# 1-5 Reexpirienceing (R)
# 6-7 avoidance (A)
# 8-12 emotional numbing (N)
# 13-15 dysphoric arousal (DA)

# 16-17 Anxios arousal (AA)
pclNoNa$pclR <- pclNoNa$PCL1 + pclNoNa$PCL2 + pclNoNa$PCL3 + pclNoNa$PCL4 + pclNoNa$PCL5
pclNoNa$pclA <- pclNoNa$PCL6 + pclNoNa$PCL7
pclNoNa$pclN <- pclNoNa$PCL8 + pclNoNa$PCL9 + pclNoNa$PCL10 + pclNoNa$PCL11 + pclNoNa$PCL12
pclNoNa$pclDA <- pclNoNa$PCL13 + pclNoNa$PCL14 + pclNoNa$PCL15
pclNoNa$pclAA <- pclNoNa$PCL16 + pclNoNa$PCL17
```


```{r binomialReg}
# Binomial regression
model2 <- glm(threeMedsCensore ~ scale(pclR) + scale(pclA) + scale(pclN) + scale(pclDA) + scale(pclAA) + AGE_OCT01 + PSYCTHR + MARRIED+ sumComorbid + sumMed, data=pclNoNa, family = binomial)
summary(model2)
summ(model2, exp = TRUE, scale =TRUE, digits = 5)
plot_model(model2, show.values = TRUE, type = "est", transform = 'exp') + theme_minimal() # plotting the model
```


```{r censore30_binomial}
model3 <- glm(censore30 ~ scale(pclR) + scale(pclA) + scale(pclN) + scale(pclDA) + scale(pclAA) + AGE_OCT01 + PSYCTHR + MARRIED+ sumComorbid + sumMed, data=pclNoNa, family = binomial)
summary(model2)
summ(model3, exp = TRUE, scale =TRUE, digits = 5)
plot_model(model3)
```

```{r plot_model2_3}
plot_summs(model2, model3, scale = TRUE, exp = TRUE, model.names = c("All year", "30 days")) #, 
           # coefs = c("Age" = "AGE_OCT01", "# Psych Sessions" = "PSYCTHR","Not Married" = "MARRIED", "Comorbidity Index" = "sumComorbid", "# of Medications" = "sumMed", "PCL-R" = '"scale(pclR)"', "PCL-A" = "scale(pclA)", "PCL-N" = "scale(pclN)", "PCL-DA" = "scale(pclDA)", "PCL-AA" = "scale(pclAA)"))
```


