---
title: "Analysis for paper"
output:
  html_document:
    df_print: paged
---
This is the final analysis for the VA paper. 
All changes will be done on this one

```{r loading_data,cache=FALSE}
# load data
library(tidyverse)
library(jtools)
library(survival)
library(survminer)
vaDat <- read_csv('/home/or/Dropbox/postdoc/Yale/va_data/MED_Drop.csv')

vaDatclean <- vaDat
# remove more than 365 psychotherapies
vaDatclean$MHTOTALLc <- ifelse((vaDatclean$MHTOTALL > 365), (vaDatclean$MHTOTALLc = 365), (vaDatclean$MHTOTALLc = vaDatclean$MHTOTALL))

# set index date
a = '01/01/1960'
vaDatclean$PTSDINDr = as.Date(as.character(a), format = '%d/%m/%Y') + vaDatclean$PTSDIND
# Now take only patients who recieved medications. 
onlyMed_dat <- subset(vaDatclean, vaDatclean$PMEDRX==1)
# set number of days from index day to drop out
vaDatclean$totalDays <- as.Date(as.character(vaDatclean$MHDROPD),format = '%m/%d/%Y')  - as.Date(vaDatclean$PTSDINDr,format = '%Y/%m/%d')
head(vaDatclean)
```


```{r creating_vars, echo=FALSE}
# create a few variables before stratification
# Race
vaDatclean$race <- dplyr::case_when(vaDatclean$BLACK==1 ~ "Black",vaDatclean$WHITE==1 ~ "White", vaDatclean$HISPANIC==1 ~ "Hispanic", vaDatclean$ASIAN ==1 ~ "Asian", TRUE ~ "Other")
vaDatclean$race <- factor(vaDatclean$race)
# change levels white will be first
vaDatclean$race <- relevel(vaDatclean$race, "White")

# turn age to categorical
vaDatclean$ageCat <- cut(vaDatclean$AGE_OCT01, breaks = c(15, 25, 35, 45, 60, 90), labels = c("18-25", "26-35", "36-45", "45-60", "60+"))
plot(vaDatclean$ageCat)
# check comorbidity
vaDatclean$sumComorbid <- vaDatclean$DEMENT_ALZ + vaDatclean$ALC + vaDatclean$DRUG + vaDatclean$SCHIZOPHRENIA + vaDatclean$OTHERPSYCHOSIS + vaDatclean$BIPOLAR + vaDatclean$MAJORAFFECT + vaDatclean$OTHERDEP + vaDatclean$ANXIETY + vaDatclean$ADJUSTMENT + vaDatclean$PERSONALITY + vaDatclean$BORDERLINE


# vaDatclean$comorbid <- dplyr::case_when(vaDatclean$DEMENT_ALZ !=0 ~ "Dementia", vaDatclean$ALC !=0 ~"Alcohol", vaDatclean$DRUG != 0 ~ "Drug", vaDatclean$SCHIZOPHRENIA != 0 ~ "Schizophrenia", vaDatclean$OTHERPSYCHOSIS!= 0 ~ "OtherPsychosis", vaDatclean$BIPOLAR != 0 ~ "Bipolar", vaDatclean$MAJORAFFECT != 0 ~ "Major Affective", vaDatclean$OTHERDEP != ~ "OtherDepression", vaDatclean$ANXIETY != 0 ~ "Anxiety", vaDatclean$ADJUSTMENT != 0 ~ "Adjustment", vaDatclean$PERSONALITY != 0 ~ "Personality", vaDatclean$BORDERLINE != 0 ~"Borderline")

vaDatclean$MARRIED <- factor(vaDatclean$MARRIED)

vaDatclean$FEMALE <- factor(vaDatclean$FEMALE)
vaDatclean$VATXM <- factor(vaDatclean$VATXM) #, labels = c("Yes", "No"))
# Creating Psychotherapy 0 - none or 1 = exist
vaDatclean$PSYCTR <- dplyr::case_when(vaDatclean$INDPSY >1 | vaDatclean$GRPPSY >1 | vaDatclean$FAMPSY >1 ~ "yes", TRUE ~ "no")
vaDatclean$PSYCTR <- factor(vaDatclean$PSYCTR)
vaDatclean$PSYCTR <- relevel(vaDatclean$PSYCTR, "yes")

vaDatclean$sumComorbidfactor <- cut(vaDatclean$sumComorbid, breaks = c(-1,0,1,3,15), labels = c("0","1", "2-3","3+"))
plot(vaDatclean$sumComorbidfactor)

vaDatclean$ANXfactor <- factor(vaDatclean$ANXIETY)
vaDatclean$URBAN_RURAL <- factor(vaDatclean$URBAN_RURAL)
vaDatclean$MAJORAFFECT <- factor(vaDatclean$MAJORAFFECT)


vaDatclean$OEFOIF <- factor(vaDatclean$OEFOIF)

vaDatclean$ADRXN_true <- case_when(vaDatclean$ADRXN > 0 ~ 1, TRUE ~0)
vaDatclean$APSRXN_true <- case_when(vaDatclean$APSRXN >0 ~ 1, TRUE ~0)
vaDatclean$AXSEDHPRXN_true <- case_when(vaDatclean$AXSEDHPRXN >0 ~ 1, TRUE ~0)
vaDatclean$STIMRXN_true <- case_when(vaDatclean$STIMRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$ACVMSRXN_true <- case_when(vaDatclean$ACVMSRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$LIRXN_true <- case_when(vaDatclean$LIRXN > 0 ~ 1, TRUE ~ 0)

vaDatclean$sumMed <- vaDatclean$LIRXN_true + vaDatclean$ACVMSRXN_true + vaDatclean$STIMRXN_true + vaDatclean$AXSEDHPRXN_true + vaDatclean$APSRXN_true + vaDatclean$ADRXN_true 
vaDatclean$sumMedCat <- factor(case_when(vaDatclean$sumMed==1 ~ '1', vaDatclean$sumMed == 2 ~ "2", vaDatclean$sumMed >= 3 ~ "3+"))

# create a censore variable who survived more than 30 days
vaDatclean$MONTH1 <- case_when(vaDatclean$totalDays <= 30 ~ 1, TRUE ~ 0)
```


Lets take just medicated patients, before stratify to different groups of medication.

```{r filter_med}
# take just medicated patients. 
medicatedVa_dat <- filter(vaDatclean, PMEDRX==1)
head(medicatedVa_dat)
# take only patients with first MH visit after index date
medicatedVa_datFirstV <-  subset(medicatedVa_dat, !is.na(medicatedVa_dat$MHFUFRST))
medicatedVa_datFirstV$timeTofV <- factor(case_when(medicatedVa_datFirstV$MHFUFRST<=7 ~ "7", medicatedVa_datFirstV$MHFUFRST>7 & medicatedVa_datFirstV$MHFUFRST<=30 ~ "7-30", TRUE ~ "30+"))

medicatedVa_datFirstV$timeTofV <- relevel(medicatedVa_datFirstV$timeTofV, "30+" )
```

```{r three_meds}
# Take anti depressants, anti psychotics and anxiolytics without any prior medications. 
threeMeds <- filter(medicatedVa_datFirstV, !is.na(AXSEDHPRXN) | !is.na(ADRXN) | !is.na(APSRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX))
```

## Medicated patient's charectaristics
```{r char}
summary(threeMeds$AGE_OCT01)
sd(threeMeds$AGE_OCT01, na.rm = TRUE)
plot(threeMeds$AGE_OCT01)
# we find 7 people above the age of 100. Talk with Ilan to decide what to do. 
ageUnreason <- filter(threeMeds, threeMeds$AGE_OCT01>100)

#####Comorbidity
table(threeMeds$sumComorbidfactor)

## number of days with medication
summary(threeMeds$MEDTOTDAY)
sd(threeMeds$MEDTOTDAY)
```
## Now we begin to run the survival analysis. 
```{r survival_analysis}
# this is the basic survival object. It can be used for specific survival graphs or for Cox Proportional Hazard
surv_objectThreeMed <- Surv(time = threeMeds$MEDTOTDAY, event = threeMeds$MHDROP)
```

```{r cox_hazard}
# Cox modeling
fit.coxphThreeMeds <- coxph(surv_objectThreeMed ~ PSYCTR + MARRIED + OEFOIF +  race + timeTofV +sumComorbid + sumMed, 
                   data = threeMeds)
# Graph
ggforest(fit.coxphThreeMeds, data = threeMeds, main = "3 Medications")
```

```{r 30_days}
hist(threeMeds$MEDTOTDAY)
threeMeds_30days <- filter(threeMeds, threeMeds$MEDTOTDAY<=30)
surv_30days <- Surv(time = threeMeds_30days$MEDTOTDAY, event = threeMeds_30days$MHDROP)
summary(surv_30days)
#Cox modeling
fit.coxphThreeMeds30 <- coxph(surv_30days ~ PSYCTR + MARRIED + OEFOIF +  race + timeTofV +sumComorbid + sumMed, 
                   data = threeMeds_30days)
# Graph
ggforest(fit.coxphThreeMeds30, data = threeMeds_30days, main = "3 Medications - 30 Days")
```
## Lastly, We are going to conduct a binomial regression analysis on patients that have valid PCL scores.

```{r pcl_charac}
# characteristics of pcl sample
pclNoNa <- filter(threeMeds, !is.na(BPCLTOT))
# pcl database contains 
nrow(pclNoNa) 
```

### A total of 109,791 patients had PCL scores.

```{r}
# create five factor pcl scores
# 1-5 Reexpirienceing (R)
# 6-7 avoidance (A)
# 8-12 emotional numbing (N)
# 13-15 disphoric arousal (DA)
# 16-17 Anxios arousal (AA)
pclNoNa$pclR <- pclNoNa$PCL1 + pclNoNa$PCL2 + pclNoNa$PCL3 + pclNoNa$PCL4 + pclNoNa$PCL5
pclNoNa$pclA <- pclNoNa$PCL6 + pclNoNa$PCL7
pclNoNa$pclN <- pclNoNa$PCL8 + pclNoNa$PCL9 + pclNoNa$PCL10 + pclNoNa$PCL11 + pclNoNa$PCL12
pclNoNa$pclDA <- pclNoNa$PCL13 + pclNoNa$PCL14 + pclNoNa$PCL15
pclNoNa$pclAA <- pclNoNa$PCL16 + pclNoNa$PCL17
pclNoNa$catpclR <- case_when(pclNoNa$pclR < median(pclNoNa$pclR, na.rm = TRUE) ~ "Low Rex", TRUE ~ "High Rex")
pclNoNa$catpclA <- case_when(pclNoNa$pclA < median(pclNoNa$pclA, na.rm = TRUE)~ "Low A", TRUE ~ "High A")
pclNoNa$catpclN <- case_when(pclNoNa$pclN < median(pclNoNa$pclN, na.rm = TRUE)~ "Low N", TRUE ~ "High N")
pclNoNa$catpclDA <- case_when(pclNoNa$pclDA < median(pclNoNa$pclDA, na.rm = TRUE)~ "Low Da", TRUE ~ "High DA")
pclNoNa$catpclAA <- case_when(pclNoNa$pclAA < median(pclNoNa$pclAA, na.rm = TRUE)~ "Low AA", TRUE ~ "High AA")
table(pclNoNa$catpclR)
```


```{r}
surv_objectPCL <- Surv(time = pclNoNa$MEDTOTDAY, event = pclNoNa$MHDROP)
summary(surv_objectPCL)
```
```{r}
# do hazard analysis using PCL R category
pclNoNa$catpclR <- factor(pclNoNa$catpclR)
pclNoNa$catpclN <- factor(pclNoNa$catpclN)
pclNoNa$catpclA <- factor(pclNoNa$catpclA)
pclNoNa$catpclDA <- factor(pclNoNa$catpclDA)
pclNoNa$catpclAA <- factor(pclNoNa$catpclAA)
fit.coxphPCL <- coxph(surv_objectPCL ~  PSYCTR + MARRIED + OEFOIF +  race + timeTofV +sumComorbid + sumMed +  catpclR + catpclN + catpclA + catpclAA + catpclDA, 
                   data = pclNoNa)
ggforest(fit.coxphPCL, data = pclNoNa)
```

```{r binomialReg}
# Binomial regression

model1 <- glm(MONTH6 ~ scale(pclR) + scale(pclA) + scale(pclN) + scale(pclDA) + scale(pclAA) + AGE_OCT01 + PSYCTR + sumComorbid + sumMed, data=pclnoNa_new, family = binomial)
summary(model1)
```

We can plot it

```{r PlotBinom}
summ(model1, exp = TRUE, scale =TRUE, digits = 5)
library(ggstance)
plot_summs(model1, scale = TRUE, plot.distributions = TRUE)
export_summs(model1, scale = TRUE, error_format = "[{conf.low}, {conf.high}]", digits = 5)
```

## check prediction of binomial regression

```{r}
pclnoNa_new <- subset(pclNoNa, !is.na(pclNoNa$pclAA) & !is.na(pclNoNa$pclA) & !is.na(pclNoNa$pclR) & !is.na(pclNoNa$pclN) & !is.na(pclNoNa$pclDA) & !is.na(pclNoNa$PSYCTR) & !is.na(pclNoNa$AGE_OCT01) & !is.na(pclNoNa$sumComorbid) & !is.na(pclNoNa$sumMed))
glm.probs = predict(model1, type = "response", na.action = rm.na)
glm.pred = ifelse(glm.probs > 0.95, "0", "1")
table(glm.pred, pclnoNa_new$MONTH6)

mean(glm.pred == pclnoNa_new$MONTH6)
```


