---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Start by cleaning the data. Taking only ani depressants.
Removing all use of medication prior to index day.
Then do hazard analysis with age (make categorical, gender, marital statues, comorbidity)

```{r}
# load data
library(tidyverse)
library(jtools)
vaDat <- read_csv('/home/or/Dropbox/postdoc/Yale/va_data/MED_Drop.csv')

vaDatclean <- vaDat
# remove more than 365 psychotherapies
vaDatclean$MHTOTALLc <- ifelse((vaDatclean$MHTOTALL > 365), (vaDatclean$MHTOTALLc = 365), (vaDatclean$MHTOTALLc = vaDatclean$MHTOTALL))

# set index date
a = '01/01/1960'
vaDatclean$PTSDINDr = as.Date(as.character(a), format = '%d/%m/%Y') + vaDatclean$PTSDIND
# Now take only patients who recieved medications. 
onlyMed_dat <- subset(vaDatclean, vaDatclean$PMEDRX==1)
# set number of days from index day to drop out
vaDatclean$totalDays <- as.Date(as.character(vaDatclean$MHDROPD),format = '%m/%d/%Y')  - as.Date(vaDatclean$PTSDINDr,format = '%Y/%m/%d')
head(vaDatclean)
```
```{r}
# create a few variables before stratification
# Race
vaDatclean$race <- dplyr::case_when(vaDatclean$BLACK==1 ~ "Black",vaDatclean$WHITE==1 ~ "White", vaDatclean$HISPANIC==1 ~ "Hispanic", vaDatclean$ASIAN ==1 ~ "Asian", TRUE ~ "Other")
vaDatclean$race <- factor(vaDatclean$race)
# change levels white will be first
vaDatclean$race <- relevel(vaDatclean$race, "White")

# turn age to categorical
vaDatclean$ageCat <- cut(vaDatclean$AGE_OCT01, breaks = c(15, 25, 35, 45, 60, 90), labels = c("18-25", "26-35", "36-45", "45-60", "60+"))
plot(vaDatclean$ageCat)
# check comorbidity
vaDatclean$sumComorbid <- vaDatclean$DEMENT_ALZ + vaDatclean$ALC + vaDatclean$DRUG + vaDatclean$SCHIZOPHRENIA + vaDatclean$OTHERPSYCHOSIS + vaDatclean$BIPOLAR + vaDatclean$MAJORAFFECT + vaDatclean$OTHERDEP + vaDatclean$ANXIETY + vaDatclean$ADJUSTMENT + vaDatclean$PERSONALITY + vaDatclean$BORDERLINE
#antiDP_dat$comorbid <- dplyr::case_when(antiDP_dat$DEMENT_ALZ!=0 ~ "Dementia", antiDP_dat$ALC !=0 ~"Alcohol", )
vaDatclean$MARRIED <- factor(vaDatclean$MARRIED)

vaDatclean$FEMALE <- factor(vaDatclean$FEMALE)
vaDatclean$VATXM <- factor(vaDatclean$VATXM) #, labels = c("Yes", "No"))
# Creating Psychotherapy 0 - none or 1 = exist
vaDatclean$PSYCTR <- dplyr::case_when(vaDatclean$INDPSY >1 | vaDatclean$GRPPSY >1 | vaDatclean$FAMPSY >1 ~ "yes", TRUE ~ "no")
vaDatclean$PSYCTR <- factor(vaDatclean$PSYCTR)
vaDatclean$PSYCTR <- relevel(vaDatclean$PSYCTR, "yes")

vaDatclean$sumComorbidfactor <- cut(vaDatclean$sumComorbid, breaks = c(-1,0,1,3,15), labels = c("0","1", "2-3","3+"))
plot(vaDatclean$sumComorbidfactor)

vaDatclean$ANXfactor <- factor(vaDatclean$ANXIETY)
vaDatclean$URBAN_RURAL <- factor(vaDatclean$URBAN_RURAL)
vaDatclean$MAJORAFFECT <- factor(vaDatclean$MAJORAFFECT)


vaDatclean$OEFOIF <- factor(vaDatclean$OEFOIF)

vaDatclean$ADRXN_true <- case_when(vaDatclean$ADRXN > 0 ~ 1, TRUE ~0)
vaDatclean$APSRXN_true <- case_when(vaDatclean$APSRXN >0 ~ 1, TRUE ~0)
vaDatclean$AXSEDHPRXN_true <- case_when(vaDatclean$AXSEDHPRXN >0 ~ 1, TRUE ~0)
vaDatclean$STIMRXN_true <- case_when(vaDatclean$STIMRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$ACVMSRXN_true <- case_when(vaDatclean$ACVMSRXN > 0 ~ 1, TRUE ~ 0)
vaDatclean$LIRXN_true <- case_when(vaDatclean$LIRXN > 0 ~ 1, TRUE ~ 0)

vaDatclean$sumMed <- vaDatclean$LIRXN_true + vaDatclean$ACVMSRXN_true + vaDatclean$STIMRXN_true + vaDatclean$AXSEDHPRXN_true + vaDatclean$APSRXN_true + vaDatclean$ADRXN_true 
vaDatclean$sumMedCat <- factor(case_when(vaDatclean$sumMed==1 ~ '1', vaDatclean$sumMed == 2 ~ "2", vaDatclean$sumMed >= 3 ~ "3+"))

# create a censore variable who survived more than 30 days
vaDatclean$MONTH1 <- case_when(vaDatclean$totalDays <= 30 ~ 1, TRUE ~ 0)

```

Lets take just medicated patients, before stratify to different groups of medication.
```{r}
# take just medicated patients. 
medicatedVa_dat <- filter(vaDatclean, PMEDRX==1)
head(medicatedVa_dat)
```
Lets look at some descriptives
```{r}
# plot age
ggplot(medicatedVa_dat, aes(x=medicatedVa_dat$AGE_OCT01)) + geom_histogram() + xlab("Age") + stat_bin(bins=30) + theme_minimal() + ggtitle("Age Histogram")
summary(medicatedVa_dat$MARRIED)
table(medicatedVa_dat$FEMALE)
plot(medicatedVa_dat$race)
g <- data.frame(table(medicatedVa_dat$race))
plot(g)
table(g)
ggplot(g, aes(x= "",y=Freq, fill = Var1)) + geom_bar(width = 14, stat = "identity") + coord_polar("y") + theme_minimal()
table(medicatedVa_dat$sumMed)
```


```{r}
# now we clean data
# take only anti depressant without any medication before index day
antiDP_dat <- dplyr::filter(vaDatclean, !is.na(ADRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 
head(antiDP_dat)

# just to check how much people using anti depressants.
antiDP <- filter(vaDatclean, !is.na(ADRXN))
294224/330772
```

Now lets look at the drugs

```{r}
# whats the average drug usage?
drugeUsage<- table(medicatedVa_dat$ADDAY)#, antiDP_dat$APSDAY, antiDP_dat$AXDAY, vaDatclean$STIMDAY)
hist(medicatedVa_dat$MEDTOTDAY)
# many patients drop out fast.
# lets see how many we have after 30 days of use
drugeMore30 <- filter(medicatedVa_dat, MEDTOTDAY >= 30)
247083/330772
# only 74.6% of the people survive the first 30 days. 
hist(drugeMore30$MEDTOTDAY)

drugLess30 <- filter(medicatedVa_dat, MEDTOTDAY  <= 30)
mean(medicatedVa_dat$ADDAY, na.rm=TRUE)
sd(antiDP_dat$ADDAY)
median(medicatedVa_dat$ADDAY, na.rm=TRUE)
plot(density(medicatedVa_dat$ADDAY, na.rm=TRUE) )
```

So - after 30 days we see major drop out. We will start by focusing on that time period. 
```{r}
# first create survival model
# drugless30 includes 84,575
84575/330772
library(survival)
library(survminer)
drugLess30$psycOne <- factor(case_when(drugLess30$INDPSY >1 | drugLess30$GRPPSY >1 | drugLess30$FAMPSY >1 ~ "yes", TRUE ~ "no"))
drugLess30$psycOne <- relevel(drugLess30$psycOne, "yes")
surv_less30 <- Surv(time = drugLess30$MEDTOTDAY, event = drugLess30$MONTH1)  
summary(surv_less30)
summary(drugLess30$MEDTOTDAY)
# now cox analysis
fit.coxphLess30 <- coxph(surv_less30 ~ psycOne + MARRIED + OEFOIF + race , 
                   data = drugLess30)
ggforest(fit.coxphLess30, data = drugLess30)
```
```{r}
# lets do survival analysis just with psychotherapy
fit1 <- survfit(surv_less30 ~ drugLess30$psycOne, data = drugLess30) # fit the analysis with gender
ggsurvplot(fit1, data = drugLess30, pval = TRUE, conf.int = TRUE)
table(drugLess30$psycOne, drugLess30$MEDTOTDAY)
mean(drugLess30$MEDTOTDAY[drugLess30$psycOne=='yes'])
```

Now we will create the same analysis, but for all patients under medication, and we will check the effect of more than one psychotherapy sessions (not more than 0)
```{r}
# 
surv_all <- Surv(time = medicatedVa_dat$MEDTOTDAY, event = medicatedVa_dat$MHDROP)  
summary(surv_all)
medicatedVa_dat$psycOne <- factor(case_when(medicatedVa_dat$INDPSY >1 | medicatedVa_dat$GRPPSY >1 | medicatedVa_dat$FAMPSY >1 ~ "yes", TRUE ~ "no"))
medicatedVa_dat$psycOne <- relevel(medicatedVa_dat$psycOne, "yes")
# now cox analysis
fit.coxAll <- coxph(surv_all ~ PSYCTR + MARRIED + OEFOIF + race + sumMedCat + sumComorbidfactor, 
                   data = medicatedVa_dat)
ggforest(fit.coxAll, data = medicatedVa_dat)
```

Now lets see the effect of psychotherapy.
```{r}
fit_psych <- survfit(surv_all ~ medicatedVa_dat$PSYCTR, data = medicatedVa_dat) # fit the analysis with gender
#summary(fit1)
ggsurvplot(fit_psych, data = medicatedVa_dat, pval = TRUE, conf.int = TRUE)
```


```{r}
model30 <- glm(MONTH1 ~ PSYCTHR + MARRIED + AGE_OCT01 + race + sumComorbid + sumMed, data = medicatedVa_dat)
summary(model30)
summ(model30, exp = TRUE, scale =TRUE, digits = 5)
library(ggstance)
plot_summs(model30, scale = TRUE, plot.distributions = TRUE)
export_summs(model30, scale = TRUE, error_format = "[{conf.low}, {conf.high}]", digits = 5)
```

Now we will omit those who didn't have first visit at all and look at the data with time to first visit. 
```{r}
medicatedVa_datFirstV <-  subset(medicatedVa_dat, !is.na(medicatedVa_dat$MHFUFRST))
307798/330772
medicatedVa_datFirstV$timeTofV <- factor(case_when(medicatedVa_datFirstV$MHFUFRST<=7 ~ "7", medicatedVa_datFirstV$MHFUFRST > 7 & medicatedVa_datFirstV$MHFUFRST <= 14 ~ "7-14", medicatedVa_datFirstV$MHFUFRST>14 & medicatedVa_datFirstV$MHFUFRST<=30 ~ "15-30", TRUE ~ "30+"))
medicatedVa_datFirstV$timeTofV <- relevel(medicatedVa_datFirstV$timeTofV, "30+")
plot(medicatedVa_datFirstV$timeTofV)
summary(medicatedVa_datFirstV$MHFUFRST)
head(medicatedVa_datFirstV)
surv_less30firstV <- Surv(time = medicatedVa_datFirstV$MEDTOTDAY, event = medicatedVa_datFirstV$MONTH1)  
summary(surv_less30firstV)
```
We can create a Cox Hazard analysis with those variables
```{r}
fit.coxphLess30FirstV <- coxph(surv_less30firstV ~ MARRIED + OEFOIF
                    + race + sumComorbidfactor + sumMedCat, 
                   data = medicatedVa_datFirstV)
ggforest(fit.coxphLess30FirstV, data = medicatedVa_datFirstV)
```
##H2

Lets look at first visit not only in 30 days. 

```{r}
surv_firstV <- Surv(time = medicatedVa_datFirstV$MEDTOTDAY, event = medicatedVa_datFirstV$MHDROP)  
summary(surv_firstV)
fit.coxphFirstV <- coxph(surv_firstV ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbidfactor + sumMedCat + timeTofV, 
                   data = medicatedVa_datFirstV)
ggforest(fit.coxphFirstV, data = medicatedVa_datFirstV)
```


Now we go to antidepressants:

```{r}
151131/330772
plot(antiDP_dat$sumMedCat)
table(antiDP_dat$sumMedCat)
plot(antiDP_dat$ageCat)

```

Now lets do survival analysis

```{r}
surv_object <- Surv(time = antiDP_dat$MEDTOTDAY, event = antiDP_dat$MHDROP)
summary(surv_object)
```


```{r}
# fitting
fit1 <- survfit(surv_object ~ antiDP_dat$VATXM, data = antiDP_dat) # fit the analysis with gender
#summary(fit1)
ggsurvplot(fit1, data = antiDP_dat, pval = TRUE, conf.int = TRUE)
```


```{r}
######################################################################
######################################################################
############### ANALYSIS ANTI DEPRESSANTS #################
fit.coxph3 <- coxph(surv_object ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbidfactor + sumMedCat, 
                   data = antiDP_dat)
ggforest(fit.coxph3, data = antiDP_dat, main = "Anti Depressant")
```
Lets take anxiolitics now
```{r}
antiANX_dat <- dplyr::filter(vaDatclean, !is.na(AXSEDHPRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 
60641/330772
surv_objectAnx <- Surv(time = antiANX_dat$MEDTOTDAY, event = antiANX_dat$MHDROP)
fit.coxph4 <- coxph(surv_objectAnx ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbidfactor + sumMedCat, 
                   data = antiANX_dat)
summary(fit.coxph4)
ggforest(fit.coxph4, data = antiANX_dat)
```

Anti psychotics
```{r}
antiAPX_dat <- dplyr::filter(vaDatclean, !is.na(APSRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX)) 

28997/330772
surv_objectApx <- Surv(time = antiAPX_dat$MEDTOTDAY, event = antiAPX_dat$MHDROP)
fit.coxph5 <- coxph(surv_objectApx ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbidfactor + sumMedCat, 
                   data = antiAPX_dat)
ggforest(fit.coxph5, data = antiAPX_dat, main = "Anti Psychotic")
```


Lets check MHFUFRST. First meeting after index day
```{r}
# many NA in first visit after index day.
antiDP_dat_noMFirstNA <- subset(antiDP_dat, !is.na(antiDP_dat$MHFUFRST))
median(antiDP_dat_noMFirstNA$MHFUFRST)
# lets create a median split groups
antiDP_dat_noMFirstNA$firstV <- factor(case_when(antiDP_dat_noMFirstNA$MHFUFRST<=30 ~ "Low_Waiting", antiDP_dat_noMFirstNA$MHFUFRST>30 ~ "High_Waiting"))
plot(antiDP_dat_noMFirstNA$firstV)
```

Now we will create a new analysis, only on this data set. 
```{r}
surv_object_clean <- Surv(time = antiDP_dat_noMFirstNA$MEDTOTDAY, event = antiDP_dat_noMFirstNA$MHDROP)
summary(surv_object)
fitClean1<-survfit(surv_object_clean ~ antiDP_dat_noMFirstNA$firstV, data = antiDP_dat_noMFirstNA) # fit the analysis with gender
#summary(fit1)
ggsurvplot(fitClean1, data = antiDP_dat_noMFirstNA, pval = TRUE)
```

```{r}
antiDP_dat_noMFirstNA$firstV <- factor(antiDP_dat_noMFirstNA$firstV)
fit.coxphClean <- coxph(surv_object_clean ~ FEMALE  + VATXM + MARRIED + ageCat + race + firstV, 
                   data = antiDP_dat_noMFirstNA)
ggforest(fit.coxphClean, data = antiDP_dat_noMFirstNA)
```

#H1
Looking at effect of symptom clusters on adherence. 

```{r}
pclNoNa <- filter(antiDP_dat, !is.na(BPCLTOT))
head(pclNoNa)

```


```{r}
# create five factor pcl scores
# 1-5 Reexpirienceing (R)
# 6-7 avoidance (A)
# 8-12 emotional numbing (N)
# 13-15 disphoric arousal (DA)
# 16-17 Anxios arousal (AA)
pclNoNa$pclR <- pclNoNa$PCL1 + pclNoNa$PCL2 + pclNoNa$PCL3 + pclNoNa$PCL4 + pclNoNa$PCL5
pclNoNa$pclA <- pclNoNa$PCL6 + pclNoNa$PCL7
pclNoNa$pclN <- pclNoNa$PCL8 + pclNoNa$PCL9 + pclNoNa$PCL10 + pclNoNa$PCL11 + pclNoNa$PCL12
pclNoNa$pclDA <- pclNoNa$PCL13 + pclNoNa$PCL14 + pclNoNa$PCL15
pclNoNa$pclAA <- pclNoNa$PCL16 + pclNoNa$PCL17
pclNoNa$catpclR <- case_when(pclNoNa$pclR < median(pclNoNa$pclR, na.rm = TRUE) ~ "Low Rex", TRUE ~ "High Rex")
pclNoNa$catpclA <- case_when(pclNoNa$pclA < median(pclNoNa$pclA, na.rm = TRUE)~ "Low A", TRUE ~ "High A")
pclNoNa$catpclN <- case_when(pclNoNa$pclN < median(pclNoNa$pclN, na.rm = TRUE)~ "Low N", TRUE ~ "High N")
pclNoNa$catpclDA <- case_when(pclNoNa$pclDA < median(pclNoNa$pclDA, na.rm = TRUE)~ "Low Da", TRUE ~ "High DA")
pclNoNa$catpclAA <- case_when(pclNoNa$pclAA < median(pclNoNa$pclAA, na.rm = TRUE)~ "Low AA", TRUE ~ "High AA")
table(pclNoNa$catpclR)
```


```{r}
ggplot(pclNoNa, aes(x=pclNoNa$pclR, y= pclNoNa$MEDTOTDAY)) + geom_smooth() 
cor.test(pclNoNa$pclAA, pclNoNa$MEDTOTDAY)
```
```{r}
surv_objectPCL <- Surv(time = pclNoNa$MEDTOTDAY, event = pclNoNa$MHDROP)
summary(surv_objectPCL)
```
```{r}
# do hazard analysis using PCL R category
pclNoNa$catpclR <- factor(pclNoNa$catpclR)
pclNoNa$catpclN <- factor(pclNoNa$catpclN)
pclNoNa$catpclA <- factor(pclNoNa$catpclA)
pclNoNa$catpclDA <- factor(pclNoNa$catpclDA)
pclNoNa$catpclAA <- factor(pclNoNa$catpclAA)
fit.coxphPCL <- coxph(surv_objectPCL ~ FEMALE  + VATXM + MARRIED + ageCat + race  + MAJORAFFECT + catpclR + catpclN + catpclA + catpclAA + catpclDA, 
                   data = pclNoNa)
ggforest(fit.coxphPCL, data = pclNoNa)
```

```{r}
# Binomial regression

model1 <- glm(MONTH2 ~ scale(pclR) + scale(pclA) + scale(pclN) + scale(pclDA) + scale(pclAA) + AGE_OCT01 + VATXM, data=pclNoNa)
summary(model1)
```

Lets look only at women:
```{r}
# just women
onlyWomen <- filter(medicatedVa_dat, FEMALE==1)
surv_objectWomen <- Surv(time = onlyWomen$MEDTOTDAY, event = onlyWomen$MHDROP)
fit.coxphWomen <- coxph(surv_objectWomen ~ PSYCTR + MARRIED + OEFOIF + race + sumComorbidfactor + sumMedCat, 
                   data = onlyWomen)
ggforest(fit.coxphWomen, data = onlyWomen, main = "Only Woman")
```
#H1
Now we will run a new analysis. Taking into account only 3 kinds of medications: anti depressants, anxiolytics and anti psychotics

```{r}
# Take anti depressants, anti psychotics and anxiolytics without any prior medications. 
threeMeds <- filter(medicatedVa_datFirstV, !is.na(AXSEDHPRXN) | !is.na(ADRXN) | !is.na(APSRXN) & is.na(PREADRX) & is.na(PREAPRX) & is.na(PREAXRX) & is.na(PRESTIRX) & is.na(PREACRX) & is.na(PRELIRX) & is.na(PREOPRX) & is.na(PREPRZRX) )
surv_objectThreeMed <- Surv(time = threeMeds$MEDTOTDAY, event = threeMeds$MHDROP)
```

A total of 318,738 patients received either AD, ANX or APX without previous medications (prior to index day).
A total of 279,809 also had one more MH meeting after index day

```{r}
fit.coxphThreeMeds <- coxph(surv_objectThreeMed ~ PSYCTR + MARRIED + OEFOIF +  race + sumComorbidfactor + sumMedCat, 
                   data = threeMeds)
ggforest(fit.coxphThreeMeds, data = threeMeds, main = "3 Medications")
```

```{r}
library(DiagrammeR)
grViz(
  "digraph rmarkdown {
node [shape = box, width = 2.5, fillcolor = lightCyan, style = filled]
a [label = 'Total Medicated\n 330,721', fillcolor = Azure] 
b [label = 'Three medications\n 318,738']
c [label = 'Three Medications and visit\n 279,809', fillcolor = cyan]
d [label = 'Total number of patients\n 502,036', fillcolor = white]
a -> b [label = 'Including AD, AX and APX', fontsize = 10] 
b-> c [label = 'Including only ones with MH visit after index day', fontsize = 10]
d -> a

  }"
)
```

