---
title: "Analyses for Network approach paper"
author: "Or Duek"
output:
  html_document:
    df_print: paged
---


## This is a *Analysis* script for network approach analysis of PCL and PCL + PHQ9 networks. Cohort of VA/DoD administrative data set. 

```{r}
library(bootnet)
require(tidyverse)
require("corrplot")       ## correlation matrix plots
require(qgraph)
require(psychonetrics)
# load data set
source('/home/or/Documents/va_data/readData.r')
```

## Next we do some cleanning and then  descriptive statistics:

```{r}
# Addind time difference between PCL and PHQ
# gather info on both meds and no meda

# all patientes with PTSD and PCLTOT
pclAll <- dplyr::filter(vaDatclean, !is.na(BPCLTOT))
# plot pcl total score 
hist(pclAll$BPCLTOT)
# we have a minimum of 2 - so we have some NAs - let remove them
pclAll_Nas <- filter(pclAll, BPCLTOT <=16)
# total of 20 subjects with 16 or less in PCL (i.e. at least one missing variable)
# we can remove them from analysis
pclAll <- filter(pclAll, BPCLTOT >=17)
# 159577 patients
#pclNetwork <- pclNoNa # just medicated
pclNetwork <- pclAll
nrow(pclNetwork)
hist(pclNetwork$BPCLTOT)

pclNetwork$PCL_PHQdiff <- pclNetwork$PHQSURVEYDATE - pclNetwork$PCLSURVEYDATE
pclPHQNetwork <- filter(pclNetwork, PCL_PHQdiff <= 14 & PCL_PHQdiff >= 0) # removing patients with more than 14 apart between PHQ9 and PCL-M
pclPHQNetwork <- filter(pclPHQNetwork, BPCLTOT>=17)
hist(pclPHQNetwork$PCL_PHQdiff)
```

## Do sample descriptives on all subjects
```{r sample descriptives}
# gather info on both meds and no meda
# remove patients with more than 14 days apart PHQ and PCL
summary(pclAll$AGE_OCT01)
mean(pclAll$AGE_OCT01, na.rm=TRUE)
sd(pclAll$AGE_OCT01, na.rm=TRUE)
summary(pclAll$BPCLTOT)
mean(pclAll$BPCLTOT)
sd(pclAll$BPCLTOT)
table(pclAll$FEMALE)
summary(pclPHQNetwork$PHQSUMTOTAL)
mean(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
sd(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
table(pclPHQNetwork$FEMALE)

```

```{r}
# build data set only with PCL items
pclItems <- dplyr::select(pclAll, starts_with("PCL"))
pclPHQItems <- dplyr::select(pclPHQNetwork, starts_with("PCL"))

pclItems_noCluster <- dplyr::select(pclItems, -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE)
nrow(pclItems_noCluster)
pclItems_noCluster <- na.omit(pclItems_noCluster)
nrow(pclItems_noCluster)


pclPHQNetwork_noCluster <- dplyr::select(pclPHQNetwork, starts_with("PCL"), -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE, -PCL_PHQdiff, PHQ1,PHQ2,PHQ3, PHQ4,PHQ5,PHQ6,PHQ7,PHQ8,PHQ9)
pclPHQNetwork_noCluster <- na.omit(pclPHQNetwork_noCluster)

```

## We compare three models to set the one's we would use thorugh the rest of analysis
### First we do it on the PCL items alone (whole 150k)
```{r}
### A. Gaussian Graphical Model, regularized
df2 <- pclItems_noCluster
n2 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2 <- plot(n2, legend.cex=.5, vsize=7)
# Severely skewed data, so we use Spearman over polychoric correlations here, as recommended (https://psycnet.apa.org/record/2018-13501-001)
# Warning: Dense network selected. But no negative edges, and bootstrapped edge weights look OK.
# However, we can still try threshold=TRUE as recommended; see next.

### B. Gaussian Graphical Model, regularized & thresholded
n3 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=TRUE)
g3 <- plot(n3, layout=g2$layout, legend.cex=.5, vsize=7)


### C. Robustness: use new estimation procedure ggmModSelect (http://psychosystems.org/qgraph_1.5)
n4 <- estimateNetwork(df2, default="ggmModSelect", corMethod = "cor", corArgs = list(method="spearman"))
g4 <- plot(n4, layout=g2$layout, legend.cex=.5, vsize=7)

## Nonregularized network - I've used this one simply enough - please comment on better methods
labels <- names(df2)

library(GGMnonreg)
n5 <- GGMregress(df2,IC = "BIC", method= "forward")
g5 <- qgraph(n5$pcor_or, layout = g2$layout, theme = "colorblind", labels = labels)


## Using Bayesian 
library(BGGM)
n6 <- explore(df2, prior_sd = 0.5, iter = 5000, cores = 4) #prior_sd: 0.5 is standard and conservative; iter= number of iterations; cores = number of CPU cores
# H1 for this network is, that edges are not zero (either negative or positive), H0 is that they are zero #
E <- select(n6, BF_cut = 10, alternative = "two.sided") # two.sided = egdes can be positive or negative
## Summary table of the network ##
summary(E, summarize = T, log = TRUE, digits = 2)  # log TRUE: BF= ln(BF), BF_10 = evidence for H1, BF_01 = evidence for H0

g6 <- qgraph(E$partials_non_zero,
       legend = FALSE, layout = g2$layout, theme = "colorblind", labels = labels)


### Bayesian R2 ###
# Compute Bayes R2 #
fit_R2 <- estimate(df2, iter = 5000)
# The following code gives me an error: no applicable method for 'predict' applied to an object of class "estimate". Tobias, do you have any idea?
# R2 <- predict(fit_R2,
#               ci_width = 0.90,
#               samples = 1000,
#               measure = "R2")

require(OpenMx)
cor(vechs(n2$graph), vechs(n4$graph)) # 0.997
cor(vechs(n2$graph), vechs(n4$graph), method="spearman") # 0.993
cor(vechs(n2$graph), vechs(n3$graph), method="spearman") #0.99
cor(vechs(n4$graph), vechs(n3$graph), method="spearman") #0.989
cor(vechs(n5$pcor_or), vechs(n2$graph), method = "spearman") #0.958
cor(vechs(E$partials_non_zero), vechs(n2$graph), method = "spearman") #0.95
# We choose n2 from here on, although looks like all are very similar - do you suggest different network to go with?
```

### Then we run the same for PCL and PHQ9 data
## Looks like all networks are highly correlated with each other. Next we will do the same test to PCL&PHQ9 network
```{r}
df3 <- pclPHQNetwork_noCluster
n2_phq <- estimateNetwork(df3, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2_phq <- plot(n2_phq, legend.cex=.5, vsize=7)


### B. Gaussian Graphical Model, regularized & thresholded
n3_phq <- estimateNetwork(df3, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=TRUE)
g3_phq <- plot(n3_phq, layout=g2_phq$layout, legend.cex=.5, vsize=7)
# Network too sparse, enforcing high specificity at the cost of sensitivity not reasonable here.

### C. Robustness: use new estimation procedure ggmModSelect (http://psychosystems.org/qgraph_1.5)
n4_phq <- estimateNetwork(df3, default="ggmModSelect", corMethod = "cor", corArgs = list(method="spearman"))
g4_phq <- plot(n4_phq, layout=g2_phq$layout, legend.cex=.5, vsize=7)


## New methods
## Nonregularized network - I've used this one simply enough - please comment on better methods
labels <- names(df3)

library(GGMnonreg)
n5_phq <- GGMregress(df3,IC = "BIC", method= "forward")
g5_phq <- qgraph(n5_phq$pcor_or, layout = g2_phq$layout, theme = "colorblind", labels = labels)


## Using Bayesian 
library(BGGM)
n6_phq <- explore(df3, prior_sd = 0.5, iter = 5000, cores = 4) #prior_sd: 0.5 is standard and conservative; iter= number of iterations; cores = number of CPU cores
# H1 for this network is, that edges are not zero (either negative or positive), H0 is that they are zero #
E_phq <- select(n6_phq, BF_cut = 10, alternative = "two.sided") # two.sided = egdes can be positive or negative
## Summary table of the network ##
summary(E_phq, summarize = T, log = TRUE, digits = 2)  # log TRUE: BF= ln(BF), BF_10 = evidence for H1, BF_01 = evidence for H0

g6_phq <- qgraph(E_phq$partials_non_zero,
       legend = FALSE, layout = g2_phq$layout, theme = "colorblind", labels = labels)



require(OpenMx)
# n3 stands out, too sparse; similarity n2 and n4 for robustness analyses:
cor(vechs(n2_phq$graph), vechs(n4_phq$graph)) # 0.995
cor(vechs(n2_phq$graph), vechs(n4_phq$graph), method="spearman") # 0.973
cor(vechs(n2_phq$graph), vechs(n3_phq$graph), method="spearman") #0.92
cor(vechs(n4_phq$graph), vechs(n3_phq$graph), method="spearman") #0.94
cor(vechs(n5_phq$pcor_or), vechs(n2_phq$graph), method = "spearman") #0.958
cor(vechs(E_phq$partials_non_zero), vechs(n2_phq$graph), method = "spearman") #0.948
# We choose n2_phq from here on - should we go with a different one? 

# run goldbricker to test correlation between nodes
library(networktools)
gold <- goldbricker(df3)
print(gold$suggested_reductions)
print(gold$proportion_matrix)
# If I'm doing it right - no suggested reductions
# 
# just want to see wat's going on
cor.test(df3$PCL13, df3$PHQ3) # highly correlated

cor.test(df3$PCL15, df3$PHQ7)
```
## OK -  so we will use n2 for PCL and n2_phq for PCL and PHQ9 analysis.

Should consider running other kinds of analysis:
1. Only PCL - but remove all participants that will take part in PCL&PHQ one
2. Only PCL - but only with participants that also has PHQ data (aroun 30k)
3. Remove item 3 from PHQ9, as it asks about sleep, similar to item 13 in PCL

### Now we run simple network analysis on PCL scores, check stability. Then do the same on PCL+PHQ score

```{r network_pcl}
# building two kinds of group clusters. One (gr1) taken from Harpaz-Rotem, I., Tsai, J., Pietrzak, R. H., & Hoff, R. (2014).
gr1 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17)) #PTSD Clusters
gr_likeDSM <- list("Intrusion"=c(1:5), "Avoidance"=c(6:12), "Arousal"=c(13:17)) #PTSD symptoms categories B C D

g2 <- plot(n2, legend.cex=.5, vsize=7, theme = "colorblind", groups = gr1)
```
# Centrality graph for this network:
```{r}
library(qgraph)
# sort by level of centrality (strength)
sort(centrality(n2)$OutDegree, decreasing = T)
# plot centrality
centralityPlot(n2, include = "all")
```

Check stability of network
```{r}
# Bootstrap 1:
#boot1 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "nonparametric") 
#plot(boot1, labels = F, order = "sample")  + theme_minimal()
# now lets look at subjects
#boot2 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "case")
#plot(boot2, labels = F, order = "sample") + theme_minimal()

# Measure Centrality stability
#boot3 <- bootnet(n2, nCores = 6, nBoots = 1000, statistics = "strength", type = "case")
#plot(boot3, labels = F, order = "sample") + theme_minimal()

#plot sig diff edges
#plot(boot1, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")

#plot sig diff nodes
#plot(boot1, "Strength") 
```

## Running same analysis on PCL with PHQ9 data

```{r}
grep("PHQ", colnames(pclPHQNetwork_noCluster)) # get location of PHQ
gr2 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17), "PHQ9-Depression"=c(18:26)) #PTSD symptoms categories B C D E  
plot(n2_phq, groups = gr2, theme="colorblind")
centralityPlot(n2_phq, include = "all")
```
# run stability testings
```{r}
# Bootstrap 1:
boot1_phq <- bootnet(n2_phq, nCores = 6, nBoots = 1000, type = "nonparametric") 
plot(boot1_phq, labels = F, order = "sample")  + theme_minimal()
# now lets look at subjects
boot2_phq <- bootnet(n2_phq, nCores = 6, nBoots = 1000, type = "case")
plot(boot2_phq, labels = F, order = "sample") + theme_minimal()

# Measure Centrality stability
#boot3 <- bootnet(n2, nCores = 6, nBoots = 1000, statistics = "strength", type = "case")
#plot(boot3, labels = F, order = "sample") + theme_minimal()

#plot sig diff edges
plot(boot1_phq, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")

#plot sig diff nodes
plot(boot1_phq, "Strength") 
```


# present a graph with centrality of both networks
```{r}
centralityPlot(list("Only PCL" = n2, "PCL_PHQ9" = n2_phq), include = c("Betweenness", "Strength", "Closeness", "ExpectedInfluence"), decreasing = TRUE) + theme_minimal() 

# if we graph only the PCL 17 
#centralityPlot(list("Only PCL" = results_noCluster, "PCL_PHQ9" = results_PHQnoCluster$graph[1:17,1:17], "onlyPCL_PHQ_data" = results_onlyPCL), include = c("Betweenness", "Strength", "Closeness", "ExpectedInfluence"), decreasing = TRUE) + theme_minimal() 
```

## Next we will run confirmatory analysis. We start with the PCL only data. We randomly sample 50% of population and run the model, then fit on the other 50%

# Split Data into Training and Testing in R 
```{r}
sample_size = floor(0.5*nrow(pclItems_noCluster))
set.seed(777)

# randomly split data in r
picked = sample(seq_len(nrow(pclItems_noCluster)),size = sample_size)
train =pclItems_noCluster[picked,]
test =pclItems_noCluster[-picked,]

# Start run confirmatory analysis
require(psychonetrics)
# run model on half the subjects 
net <- estimateNetwork(train, default = "ggmModSelect", verbose = FALSE)
network <- 1*(net$graph != 0)
model_frombootnet <- ggm(train, omega = network) %>% runmodel

```

# Run confirmatory analysis

```{r}
adjacency <- network #1*(getmatrix(model_frombootnet, "omega")!=0)
confirmatory <- ggm(test, omega = adjacency)
confirmatory <- confirmatory %>% runmodel

confirmatory %>% fit
```
```{r}
compare(train = model_frombootnet , test = confirmatory)
```
## Running confirmatory analysis on PCL and PHQ9
```{r}
pclPHQNetwork_noCluster

# Split Data into Training and Testing in R 
sample_sizePHQ = floor(0.5*nrow(pclPHQNetwork_noCluster))
set.seed(777)

# randomly split data in r
pickedPHQ = sample(seq_len(nrow(pclPHQNetwork_noCluster)),size = sample_sizePHQ)
trainPHQ =pclPHQNetwork_noCluster[pickedPHQ,]
testPHQ =pclPHQNetwork_noCluster[-pickedPHQ,]

# Start run confirmatory analysis
require(psychonetrics)
# run model on half the subjects 
netPHQ <- estimateNetwork(trainPHQ, default = "ggmModSelect", verbose = FALSE)
networkPHQ <- 1*(netPHQ$graph != 0)
model_frombootnetPHQ <- ggm(trainPHQ, omega = networkPHQ) %>% runmodel

```
```{r}
# confirmatory 
adjacencyPHQ <- networkPHQ # fixed according to Eiko's comment
confirmatoryPHQ <- ggm(testPHQ, omega = adjacencyPHQ)
confirmatoryPHQ <- confirmatoryPHQ %>% runmodel

confirmatoryPHQ %>% fit
```

```{r}
compare(train = model_frombootnetPHQ , test = confirmatoryPHQ)
```
## As suggested - I've added data driven clustering using EGA
```{r}
library(EGAnet)
# CLuter of PCL and PHQ
ega<-EGA(df3, plot.EGA = TRUE, steps = 4)

# CLuter only PCL
egaPCL<-EGA(df2, plot.EGA = TRUE, steps = 4)
# use different load
egaPCL_load0.5<-EGA(df2, plot.EGA = TRUE, steps = 4, load = 0.5)

```
## Only PCL has three clusters (when EGA load is set to 0.7 )

# Next things
1. Should deal with PCL15 and PHQ9-7 (two concentration) , and PCL13 anf PHQ9-3 (two sleep) - please see tests above and comment on that. 
2. Should we run DAG also (Tobias against, Eiko also?) - For now - removing DAG
3. 