---
title: "Analyses for Network approach paper"
author: "Or Duek"
output:
  html_document:
    df_print: paged
---


## This is a *Analysis* script for network approach analysis of PCL and PCL + PHQ9 networks. Cohort of VA/DoD administrative data set. 

```{r}
library(bootnet)
require(tidyverse)
require("corrplot")       ## correlation matrix plots
require(qgraph)
require(psychonetrics)
# load data set
source('/home/or/Documents/va_data/readData.r')
```

## Next we do some cleanning and then  descriptive statistics:

```{r}
# Addind time difference between PCL and PHQ
# gather info on both meds and no meda

# all patientes with PTSD and PCLTOT
pclAll <- dplyr::filter(vaDatclean, !is.na(BPCLTOT))
# plot pcl total score 
hist(pclAll$BPCLTOT)
# we have a minimum of 2 - so we have some NAs - let remove them
pclAll_Nas <- filter(pclAll, BPCLTOT <=16)
# total of 20 subjects with 16 or less in PCL (i.e. at least one missing variable)
# we can remove them from analysis
pclAll <- filter(pclAll, BPCLTOT >=17)
# 159577 patients
#pclNetwork <- pclNoNa # just medicated
pclNetwork <- pclAll
nrow(pclNetwork)
hist(pclNetwork$BPCLTOT)

pclNetwork$PCL_PHQdiff <- pclNetwork$PHQSURVEYDATE - pclNetwork$PCLSURVEYDATE
pclPHQNetwork <- filter(pclNetwork, PCL_PHQdiff <= 14 & PCL_PHQdiff >= 0) # removing patients with more than 14 apart between PHQ9 and PCL-M
pclPHQNetwork <- filter(pclPHQNetwork, BPCLTOT>=17)
hist(pclPHQNetwork$PCL_PHQdiff)
```

## Do sample descriptives on all subjects
```{r sample descriptives}
# gather info on both meds and no meda
# remove patients with more than 14 days apart PHQ and PCL
summary(pclAll$AGE_OCT01)
mean(pclAll$AGE_OCT01, na.rm=TRUE)
sd(pclAll$AGE_OCT01, na.rm=TRUE)
summary(pclAll$BPCLTOT)
mean(pclAll$BPCLTOT)
sd(pclAll$BPCLTOT)
table(pclAll$FEMALE)
summary(pclPHQNetwork$PHQSUMTOTAL)
mean(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
sd(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
table(pclPHQNetwork$FEMALE)

```

```{r}
# build data set only with PCL items
pclItems <- select(pclAll, starts_with("PCL"))
pclPHQItems <- dplyr::select(pclPHQNetwork, starts_with("PCL"))

pclItems_noCluster <- select(pclItems, -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE)
nrow(pclItems_noCluster)
pclItems_noCluster <- na.omit(pclItems_noCluster)
nrow(pclItems_noCluster)


pclPHQNetwork_noCluster <- select(pclPHQNetwork, starts_with("PCL"), -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE, -PCL_PHQdiff, PHQ1,PHQ2,PHQ3, PHQ4,PHQ5,PHQ6,PHQ7,PHQ8,PHQ9)
pclPHQNetwork_noCluster <- na.omit(pclPHQNetwork_noCluster)

```

## We compare three models to set the one's we would use thorugh the rest of analysis
### First we do it on the PCL items alone (whole 150k)
```{r}
### A. Gaussian Graphical Model, regularized
df2 <- pclItems_noCluster
n2 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2 <- plot(n2, legend.cex=.5, vsize=7)
# Severely skewed data, so we use Spearman over polychoric correlations here, as recommended (https://psycnet.apa.org/record/2018-13501-001)
# Warning: Dense network selected. But no negative edges, and bootstrapped edge weights look OK.
# However, we can still try threshold=TRUE as recommended; see next.

### B. Gaussian Graphical Model, regularized & thresholded
n3 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=TRUE)
g3 <- plot(n3, layout=g2$layout, legend.cex=.5, vsize=7)


### C. Robustness: use new estimation procedure ggmModSelect (http://psychosystems.org/qgraph_1.5)
n4 <- estimateNetwork(df2, default="ggmModSelect", corMethod = "cor", corArgs = list(method="spearman"))
g4 <- plot(n4, layout=g2$layout, legend.cex=.5, vsize=7)

 
require(OpenMx)

cor(vechs(n2$graph), vechs(n4$graph)) # 0.997
cor(vechs(n2$graph), vechs(n4$graph), method="spearman") # 0.993
cor(vechs(n2$graph), vechs(n3$graph), method="spearman") #0.99
cor(vechs(n4$graph), vechs(n3$graph), method="spearman") #0.989

# We choose n2 from here on, although looks like all are very similar. 
```

### Then we run the same for PCL and PHQ9 data
## Looks like all networks are highly correlated with each other. Next we will do the same test to PCL&PHQ9 network
```{r}
df3 <- pclPHQNetwork_noCluster
n2_phq <- estimateNetwork(df3, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2 <- plot(n2_phq, legend.cex=.5, vsize=7)


### B. Gaussian Graphical Model, regularized & thresholded
n3_phq <- estimateNetwork(df3, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=TRUE)
g3 <- plot(n3_phq, layout=g2$layout, legend.cex=.5, vsize=7)
# Network too sparse, enforcing high specificity at the cost of sensitivity not reasonable here.

### C. Robustness: use new estimation procedure ggmModSelect (http://psychosystems.org/qgraph_1.5)
n4_phq <- estimateNetwork(df3, default="ggmModSelect", corMethod = "cor", corArgs = list(method="spearman"))
g4 <- plot(n4_phq, layout=g2$layout, legend.cex=.5, vsize=7)

 
require(OpenMx)
# n3 stands out, too sparse; similarity n2 and n4 for robustness analyses:
cor(vechs(n2_phq$graph), vechs(n4_phq$graph)) # 0.995
cor(vechs(n2_phq$graph), vechs(n4_phq$graph), method="spearman") # 0.973
cor(vechs(n2_phq$graph), vechs(n3_phq$graph), method="spearman") #0.92
cor(vechs(n4_phq$graph), vechs(n3_phq$graph), method="spearman") #0.94

# We choose n2_phq from here on
```
## OK -  so we will use n2 for PCL and n2_phq for PCL and PHQ9 analysis.

Should consider running other kinds of analysis:
1. Only PCL - but remove all participants that will take part in PCL&PHQ one
2. Only PCL - but only with participants that also has PHQ data (aroun 30k)
3. Remove item 3 from PHQ9, as it asks about sleep, similar to item 13 in PCL

### Now we run simple network analysis on PCL scores, check stability. Then do the same on PCL+PHQ score

```{r network_pcl}
# building two kinds of group clusters. One (gr1) taken from Harpaz-Rotem, I., Tsai, J., Pietrzak, R. H., & Hoff, R. (2014).
gr1 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17)) #PTSD Clusters
gr_likeDSM <- list("Intrusion"=c(1:5), "Avoidance"=c(6:12), "Arousal"=c(13:17)) #PTSD symptoms categories B C D

g2 <- plot(n2, legend.cex=.5, vsize=7, theme = "colorblind", groups = gr1)
```
# Centrality graph for this network:
```{r}
library(qgraph)
# sort by level of centrality (strength)
sort(centrality(n2)$OutDegree, decreasing = T)
# plot centrality
centralityPlot(n2, include = "all")
```

Check stability of network
```{r}
# Bootstrap 1:
boot1 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "nonparametric") 
plot(boot1, labels = F, order = "sample")  + theme_minimal()
# now lets look at subjects
boot2 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "case")
plot(boot2, labels = F, order = "sample") + theme_minimal()

# Measure Centrality stability
boot3 <- bootnet(n2, nCores = 6, nBoots = 1000, statistics = "strength", type = "case")
plot(boot3, labels = F, order = "sample") + theme_minimal()

#plot sig diff edges
plot(boot1, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")

#plot sig diff nodes
plot(boot1, "Strength") 
```

## Running same analysis on PCL with PHQ9 data

```{r}
grep("PHQ", colnames(pclPHQNetwork_noCluster)) # get location of PHQ
gr2 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17), "PHQ9-Depression"=c(18:26)) #PTSD symptoms categories B C D E  
plot(n2_phq, groups = gr2, theme="colorblind")
centralityPlot(n2_phq, include = "all")
```
# run stability testings
```{r}
# Bootstrap 1:
boot1_phq <- bootnet(n2_phq, nCores = 6, nBoots = 1000, type = "nonparametric") 
plot(boot1_phq, labels = F, order = "sample")  + theme_minimal()
# now lets look at subjects
boot2_phq <- bootnet(n2_phq, nCores = 6, nBoots = 1000, type = "case")
plot(boot2_phq, labels = F, order = "sample") + theme_minimal()

# Measure Centrality stability
#boot3 <- bootnet(n2, nCores = 6, nBoots = 1000, statistics = "strength", type = "case")
#plot(boot3, labels = F, order = "sample") + theme_minimal()

#plot sig diff edges
plot(boot1_phq, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")

#plot sig diff nodes
plot(boot1_phq, "Strength") 
```


# present a graph with centrality of both networks
```{r}
centralityPlot(list("Only PCL" = n2, "PCL_PHQ9" = n2_phq, "onlyPCL_PHQ_data" = results_onlyPCL), include = c("Betweenness", "Strength", "Closeness", "ExpectedInfluence"), decreasing = TRUE) + theme_minimal() 

# if we graph only the PCL 17 
#centralityPlot(list("Only PCL" = results_noCluster, "PCL_PHQ9" = results_PHQnoCluster$graph[1:17,1:17], "onlyPCL_PHQ_data" = results_onlyPCL), include = c("Betweenness", "Strength", "Closeness", "ExpectedInfluence"), decreasing = TRUE) + theme_minimal() 
```

## Next we will run confirmatory analysis. We start with the PCL only data. We randomly sample 50% of population and run the model, then fit on the other 50%

# Split Data into Training and Testing in R 
```{r}
sample_size = floor(0.5*nrow(pclItems_noCluster))
set.seed(777)

# randomly split data in r
picked = sample(seq_len(nrow(pclItems_noCluster)),size = sample_size)
train =pclItems_noCluster[picked,]
test =pclItems_noCluster[-picked,]

# Start run confirmatory analysis
require(psychonetrics)
# run model on half the subjects 
net <- estimateNetwork(train, default = "ggmModSelect", verbose = FALSE)
network <- 1*(net$graph != 0)
model_frombootnet <- ggm(train, omega = network) %>% runmodel

```

# Run confirmatory analysis

```{r}
adjacency <- 1*(getmatrix(model_frombootnet, "omega")!=0)
confirmatory <- ggm(test, omega = adjacency)
confirmatory <- confirmatory %>% runmodel

confirmatory %>% fit
```
```{r}
compare(configural = model_frombootnet , metric = confirmatory)
```
## Running confirmatory analysis on PCL and PHQ9
```{r}
pclPHQNetwork_noCluster

# Split Data into Training and Testing in R 
sample_sizePHQ = floor(0.5*nrow(pclPHQNetwork_noCluster))
set.seed(777)

# randomly split data in r
pickedPHQ = sample(seq_len(nrow(pclPHQNetwork_noCluster)),size = sample_sizePHQ)
trainPHQ =pclPHQNetwork_noCluster[pickedPHQ,]
testPHQ =pclPHQNetwork_noCluster[-pickedPHQ,]

# Start run confirmatory analysis
require(psychonetrics)
# run model on half the subjects 
netPHQ <- estimateNetwork(trainPHQ, default = "ggmModSelect", verbose = FALSE)
networkPHQ <- 1*(netPHQ$graph != 0)
model_frombootnetPHQ <- ggm(trainPHQ, omega = networkPHQ) %>% runmodel

```
```{r}
# confirmatory 
adjacencyPHQ <- 1*(getmatrix(model_frombootnetPHQ, "omega")!=0)
confirmatoryPHQ <- ggm(testPHQ, omega = adjacencyPHQ)
confirmatoryPHQ <- confirmatoryPHQ %>% runmodel

confirmatoryPHQ %>% fit
```

```{r}
compare(configural = model_frombootnetPHQ , metric = confirmatoryPHQ)
```

# Next things
1. Should deal with PCL15 and PHQ9-7 (two concentration) , and PCL13 anf PHQ9-3 (two sleep)
2. Should we run DAG also (Tobias against, Eiko also?)
3. 